\chapter{Cardinality}
\begin{dfnbox}{Cardinality}{cardinality}
    \dfntxt{Cardinality} is a measure of the amount of elements in a set, denoted $\abs{A}$. We say two sets have the same cardinality if there exists a bijection between them.
\end{dfnbox}

For finite sets, we can think of cardinality as the number of elements in that set. For infinite sets, cardinality can sometimes go against our intuition. For any sets $A,B,C$:
\begin{enumerate}[noitemsep]
    \item $\abs{A} = \abs{A}$
    \item if $\abs{A} = \abs{B}$, then $\abs{B} = \abs{A}$
    \item if $\abs{A} = \abs{B}$ and $\abs{B} = \abs{C}$, then $\abs{A} = \abs{C}$.
\end{enumerate}
Hence, equality of cardinalities is an equivalence relation.
% Above is kinda self evident? maybe delete this

\begin{exbox}{Cardinality of $\N$ and $2\N$}{}
    Let $2\N \coloneq \{ 2n : n \in \N \}$ (i.e. the set of even natural numbers). Then $\abs{\N} = \abs{2\N}$.
    \tcblower
    \begin{proof}
        To show that these two sets have the same cardinality, we need to find some bijection between the sets. Let $f : \N \to 2\N$ be a function defined by $f(n) = 2n$. Note that $f$ is well-defined (i.e. is actually a function) because $f(n) \in 2 \N$ for all $n \in \N$. To prove that $f$ is a bijection, we need to prove it is both injective and surjective.
        \begin{enumerate}
            \item Let $n_1, n_2 \in \N$ such that $f(n_1) = f(n_2)$. Then $2n_1 = 2n_2$, so $n_1 = n_2$. Thus, $f$ is injective.
            \item Let $m \in 2\N$. Then $m = 2k$ for some $k \in \N$, so $m = 2k = f(k)$ for some $k \in \N$. Thus, $f$ is surjective.
        \end{enumerate}
        Therefore, $f$ is a bijection, so $\abs{\N} = \abs{2\N}$.
    \end{proof}
\end{exbox}

\begin{exbox}{Cardinality of Intervals}{}
    Let $a,b \in \R$ where $a < b$. Then $\abs{(0,1)} = \abs{(a,b)}$.
    \tcblower
    \begin{proof}
        We need to find a bijection from $(0,1)$ to $(a,b)$. We need to ``scale'' the interval $(0,1)$ to the width of $(a,b)$, then translate it to match $(a,b)$. Define $f : (0,1) \to (a,b)$ by $f(x) = a + (b-a)x$. (We need to check $f$ is well-defined). Let $x \in (0,1)$. Then $0<x<1$, so multiplying by $(b-a)$ which is positive gives $0 < (b-a)x < b-a$. Adding $a$, we get $a < a+(b-a)x < b$. Now we need to show $f$ is a bijection:
        \begin{enumerate}
            \item Let $x_1, x_2 \in (0,1)$ such that $f(x_1) = f(x_2)$. Then $a + (b-a)x_1 = a+ (b-a)x_2$. Subtracting $a$ from both sides, we get $(b-a)x_1 = (b-a)x_2$. Since $(b-a) \neq 0$, we can divide both side by $(b-a)$ to get $x_1 = x_2$.
            \item Let $y \in (a,b)$.
            \begin{notebox}
                \textbf{Scratchwork:} We want to find some $x \in (0,1)$ where $y = f(x) = a + (b-a)x$. Using some algebra to solve for $x$, we have $x = \frac{y-a}{b-a}$
            \end{notebox}
            Let $x = \frac{y-a}{b-a}$. First, we show $x \in (0,1)$:
            \begin{alignat*}{2}
                && a < y < b \\
                & \implies \quad & 0 < y-a < b-a \\
                & \implies & 0 < \frac{y-a}{b-a} < 1
            \end{alignat*}
            Thus, $x \in (0,1)$. Also:
            \[ f(x) = a + (b-a) \left( \frac{y-a}{b-a} \right) = a + (y-a) = y \]
            Thus, $f$ is surjective.
        \end{enumerate}
        Therefore, $f$ is a bijective, so $\abs{(0,1)} = \abs{(a,b)}$.
    \end{proof}
\end{exbox}

\begin{dfnbox}{Power Set}{}
    Let $A$ be a set. The \dfntxt{power set} of $A$ is the set of all subsets of $A$.
    \tcblower
    \[ \mathcal{P}(A) = \{ B : B \subseteq A \} \]
\end{dfnbox}

For example, the power set of $\{1,2,3\}$ is $\left\{\emptyset, \{1\}, \{2\}, \{3\}, \{1,2\}, \{1,3\}, \{2,3\}, \{1,2,3\} \right\}$. For any finite set with $n$ elements in it, its power set has $2^n$ elements in it.

\begin{exbox}{Cardinality of $\N$ and $\mathcal{P}(N)$}{powerset-card}
    $\abs{\N} \neq \abs{\mathcal{P}(\N)}$
    \tcblower
    \begin{proof}
        We will show that any function $f : \N \to \mathcal{P}(\N)$ cannot be surjective, and thus not bijective. Let $f : \N \to \mathcal{P}(\N)$ be any function defined by $f(n) = A_n$. Note $A_n \subseteq \N$, so $A_n \in \mathcal{P}(A)$. Now we will define a set that isn't in $f[\N]$. For each $n \in \N$, if $n \in A_n$, then $n \notin A$, and if $n \notin A_n$, then $n \in A$. More formally, $A \coloneq \{ n \in \N : n \notin A_n \}$. For all $k \in \N$, note that:
        \begin{itemize}[noitemsep]
            \item if $k \in A_k$, then $k \notin A$, so $A \neq A_k$, and
            \item if $k \notin A_k$, then $k \in A_k$, so $A \neq A_k$.
        \end{itemize}
        Hence, $A \subseteq \N$, but $f(k) \neq A$ for any $k \in \N$. Thus, $f$ is not surjective.
    \end{proof}
\end{exbox}

\begin{dfnbox}{Finite, Countably Infinite, Countable, Uncountable}{}
    Let $A$ be a set. We say $A$ is:
    \begin{itemize}[noitemsep]
        \item \dfntxt{finite} if $A \neq \emptyset$ or $\abs{A} = \abs{\{1, 2, \ldots, n\}}$ for some $n \in \N$.
        \item \dfntxt{countably infinite} if $\abs{A} = \abs{N}$.
        \item \dfntxt{countable} if $A$ is finite or countably infinite
        \item \dfntxt{uncountable} if $A$ is not countable
    \end{itemize}
\end{dfnbox}

\begin{thmbox}{$\mathcal{P}(\N)$ is uncountable.}{}
    \begin{proof}
        We know from Example \ref{ex:powerset-card} that $\mathcal{P}(\N)$ is not countably infinite. We need to show that $\mathcal{P}(\N)$ is not finite. Since $\{1\} \in \mathcal{P}(\N)$, then it cannot be empty. Suppose for contradiction $\abs{\{1,2,\ldots,n\}} = \abs{\mathcal{P}(\N)}$ for some $n \in \N$, then there exists a bijection $f : \{1,2,\ldots,n\} \to \mathcal{P}(\N)$. Define $g : \N \to \{1,2,\ldots,n\}$ by:
        \[ g(k) = \begin{cases} k, & 1\leq k \leq n \\ 1, & k > n \end{cases} \]
        Then $g$ is surjective, so $f \circ g : \N \to \mathcal{P}(\N)$ is surjective. This contradicts the fact that no such function exists (by Example \ref{ex:powerset-card}).
    \end{proof}
\end{thmbox}

Generally, there is never a bijection from a set to its power set.

\begin{notebox}
    \textbf{Intuition:} A set is countable if its elements can be ``listed'' or ``counted''. That is, for finite sets:
    \[ X = \{x_1, x_2, \ldots, x_n\} = \{x_k\}_{k=1}^{n}\]
    For infinitely countable sets:
    \[ X = \{x_1, x_2, \ldots \} = \{x_k\}_{k=1}^\infty \]
    If $X$ is finite, then there exists a bijection $f : \{1,2,\ldots,n\} \to X$ . Thus, $X = \{f(1), f(2), \ldots, f(n) \}$. If $X$ is countably infinite, then there exists a bijection $f : \N \to X$. Thus, $X = \{f(1), f(2), \ldots\}$.
\end{notebox}

\begin{thmbox}{Subsets of Countable Sets are Countable}{card-subset}
    The subset of a countable set is still countable. (i.e. a countable set cannot contain an uncountable subset).
    \tcblower
    \begin{proof}
        Let $X$ be a countable set, and let $A \subseteq X$. We will consider two cases. First, if $A$ is finite, then $A$ is countable, and we are done. Otherwise, $A$ is infinite, and hence $X$ is infinite. Then $X$ is countably infinite, so $X = \{x_1, x_2, \ldots\} = \{x_k\}_{k=1}^\infty$.
        \begin{notebox}
            \textbf{Idea:} Our set $A$ might look something like $\{x_3, x_4, x_6, \ldots\}$. We need to align these indices to $1$, $2$, $3$, and so on. We'll let $k_1 = \min\{3,4,6,\ldots\}$, let $k_2 = \min\{4,6,\ldots\}$, and so on.
        \end{notebox}
        Let $k_1 \coloneq \min\{k \in \N : x_k \in A\}$. Let $a_1 \coloneq x_{k_1}$. For all $j \in \N$ such that $j > 1$, we define $k_j \coloneq \min\{k \in \N : (x_k \in A) \land (k > k_{j-1}) \}$. Let $a_j \coloneq x_{k_j}$. Then $1 \leq k_1 < k_2 < k_3 < \ldots$, so $k_j$ approaches infinity. Let $g : \N \to A$ be a function defined by $g(j) = a_j$. We need to show that $g$ is both injective and surjective, and thus a bijection.
        \begin{itemize}
            \item Suppose that $g(j_1) = g(j_2)$ for some $j_1, j_2 \in \N$. Then $a_{j_1} = a_{j_2}$, so $x_{k_{j_1}} = x_{k_{j_2}}$. Then $k_{j_1} = k_{j_2}$, so $j_1 = j_2$. Thus, $g$ is injective.
            \item Let $a \in A$ Since $A \subseteq X$, then $a \in X$. Thus, $a = x_l$ for some $l \in \N$. Let $m \coloneq \min\{ j \in \N : k_j \geq l \}$. Since $m \in \{ j \in \N : j_k \geq l \}$, then $k_m \geq l$. Also, $m-1 \notin \{j \in \N : k_j \geq l \}$, so $k_{m-1} < l$. Now, $k_m = \min\{k \in \N : (x_k \in A) \land (k > k_{m-1})\}$. But $x_l \in A$, and $l > k_{m-1}$, so $l \in \{k \in \N : (x_k \in A) \land (k > k_{m-1}) \}$. Thus, $k_m \leq l$, because $k_m$ is the minimum of the set containing $l$. By trichotomy, $k_m = l$. Therefore:
            \[ g(m) = a_m = x_{k_m} = x_l = a \]
            So $g$ is surjective.
        \end{itemize}
        Since $g$ is a bijection, then $\abs{\N} = \abs{A}$, so $\abs{A}$ is countable.
    \end{proof}
\end{thmbox}

\begin{thmbox}{Injectivity and Cardinality}{injectivity-cardinality}
    A set $A$ is countable if and only if there exists an injective function $f : A \to \N$.
    \tcblower
    \begin{proof}
        First, suppose $A$ is a countable set. We consider two cases:
        \begin{itemize}
            \item If $A$ is countably infinite, then there exists a bijection $f : A \to \N$.
            \item If $A$ is finite, then there exists a bijection $f : A \to \{1,2,\ldots,n\}$ for some $n \in \N$. Let $g : \{1,2,\ldots,n\} \to \N$ be a function defined by $g(x) = x$ (i.e. an inclusion mapping). Then $f$ and $g$ are both injective, so $g \circ f : A \to \N $ is injective.
        \end{itemize}

        Conversely, suppose $f : A \to \N$ is an injection. Then $f[A] \subseteq \N$, so $f[A]$ is countable by Theorem \ref{thm:card-subset}. Define $g : A \to f[A]$ by $g(a) = f(a)$. Then $g$ is injective because $f$ is injective, and $g$ is surjective because $g[A] = f[A]$. Thus, $g$ is a bijection, so $\abs{A} = \abs{f[A]}$. Therefore, $A$ is countable.
    \end{proof}
\end{thmbox}

\begin{thmbox}{$\abs{\N \times \N} = \abs{\N}$}{nxn-countable}
    $\N \times \N$ is countable.
    \tcblower
    \begin{proof}
        Let $f : \N \times \N \to \N$ be a function defined by $f(n,m) = 2^n 3^m$. We now show that $f$ is bijective. To prove $f$ is injective, suppose $f(n_1, m_1) = f(n_2, n_2)$. Then $2^{n_1} 3^{m_1} = 2^{n_2} 3^{m_2}$.
        \begin{itemize}
            \item If $n_1 > n_2$, then $2^{n_1 - n_2} = 3^{m_2 - m_1}$. Since $n_1 > n_2$, we have $n_1 - n_2 > 0$, so $2^{n_1 - n_2} \in \N$. Then also $3^{m_2 - m_1} \in \N$. But $2^{n_1 - n_2}$ is even, and $3^{m_2 - m_1}$ is odd. This contradicts the fact that $2^{n_1 - n_2} = 3^{m_2 - m_1}$.
            \item If $n_2 > n_1$, then $3^{m_1 - m_2} = 2^{n_2 - n_1}$. By a similar argument, $2^{n_2 - n_1}$ is even and $3^{m_1 - m_2}$ is odd, producing the same contradiction.
            \item If $n_1 = n_2$, then $2^{n_1} = 2^{n_2}$, so cancelling gives $3^{m_1} = 3^{m_2}$. Thus, $m_1 = m_2$.
        \end{itemize}
        Hence, $(n_1, m_1) = (n_2, m_2)$, so $f$ is injective.
        By Theorem \ref{thm:injectivity-cardinality}, $\N \times \N$ is countable. Also, $\N \times \N$ is infinite, so $\abs{\N \times \N} = \abs{\N}$.
    \end{proof}
\end{thmbox}

\begin{thmbox}{Collection of Countable Sets}{}
    Suppose that for all $k \in \N$, $A_k$ is a countable set. Then $\cup_k A_k \in \N$ is countable. (i.e. a countable union of countable sets is countable)
    \tcblower
    \begin{proof}
        Let $k \in \N$. $A_k$ is countable, so $A_k$ can be listed as such:
        \begin{align*}
            A_1 &= \{a_{11}, a_{12}, a_{13}, a_{14}, \ldots\} = \{a_j\}_{j \in \N} \\
            A_2 &= \{a_{21}, a_{22}, a_{23}, a_{24}, \ldots\} \\
            &\vdots \\
            A_k &= \{a_{k1}, a_{k2}, a_{k3}, a_{k4}, \ldots\}
        \end{align*}

        \begin{notebox}
            We want to define some function $f : \bigcup_{k \in \N}A_k \to \N \times \N$ where $f(a_{kj}) = (k,j) \in \N \times \N$.
            However, we need to consider the possibility that the sets $A_k$ are not disjoint. If $a_{12} = a_{34}$, then $a(_{12}) = (1,2)$ and $f(a_{34}) = (3,4)$.
        \end{notebox}
        Given $a \in \bigcup_{k \in \N}A_k$, let $k(a) \coloneq \min\{ k \in \N : a \in A_k \}$. If $a \in A_{k(a)}$, then there is a unique $j(a) \in \N$ such that $a = a_{k(a)j(a)}$. Now define $f : \bigcup_{k \in \N} A_k \to \N \times \N$ by $f(a) = (k(a), j(a))$. We must show that $f$ is injective. Let $x,y \in \bigcup_{k \in \N}A_k$ such that $f(x) = f(y)$. That is, $(k(x),j(x)) = (k(y),j(y))$. Then $x = a_{k(x)j(x)} = a_{k(y)j(y)} = y$. By Theorem $\ref{thm:nxn-countable}$, there exists some injection $g : \N \times \N \to \N$. Hence, $g \circ f : \bigcup_{k \in \N}A_k \to \N$ is injective. By Theorem \ref{thm:injectivity-cardinality}, $\bigcup_{k \in \N} A_k$ is countable.
    \end{proof}
\end{thmbox}

This theorem shows that, in order to prove a countable union of countable sets is countable, we just need to show that each set in the union is countable. We'll use this in our proof that the set of rational numbers is a countable set.

\begin{thmbox}{$\Q$ is Countable}{}
    \begin{proof}
        Let $\Q^+ \coloneq \{r \in \Q : r > 0\}$, and let $\Q^- \coloneq \{ r \in \Q : r < 0 \}$. First, we'll prove that $\Q^+$ is countable. Let $f : \Q^+ \to \N \times \N$ be a function defined as $f(r) = (p,q)$ such that $r = \sfrac{p}{q}$ where $p,q \in \N$ and $p$ shares no common factors with $q$. To show $f$ is injective, let $r_1, r_2 \in \Q$ where  $f(r_1) = f(r_2)$. Then $r_1 = \sfrac{p_1}{q_1}, r_2 = \sfrac{p_2}{q_2}$ where $p_1, q_1 \in \N$ with no common factors, and $p_2, q_2 \in \N$ with no common factors. Thus, $(p_1,q_1) = (p_2, q_2)$, so $p_1 = p_2$ and $q_1 = q_2$. Thus, $r_1 = \sfrac{p_1}{q_1} = \sfrac{p_2}{q_2} = r_2$, so $f$ is injective. Since there exists an injection $g \in \N \times \N \to \N$, then $g \circ f : \Q^+ \to \N$ is injective. Thus, $\Q^+$ is countable.

        Next, we'll prove that $\Q^-$ is countable. Let $h : \Q^- \to \Q^+$ by $h(r) = -r$. We show $h$ is injective. If $h(r_1) = h(r_2)$ where $r_1, r_2 \in \Q^-$, then $-r_1 = -r_2$, so $r_1 = r_2$. Thus, $h$ is injective. From above, there exist an injection $\phi : \Q^+ \to \N$. Hence, $h \circ \phi : \Q^{-1} \to \N$ is injective.

        Finally, $\{0\}$ is countable because it is finite. Since $\Q = \Q^+ \cup \{0\} \cup \Q^-$ is a countable union of countable sets, then $\Q$ is countable.
    \end{proof}
\end{thmbox}

This means we can ``list'' the rational numbers (disregarding order) as $\Q = \{r_1, r_2, \ldots\} = \{r_n\}_{n \in \N}$.

\begin{thmbox}{$\R$ is Uncountable}{}
    \begin{proof}
        Suppose for contradiction $\R$ is countable. Then $\R$ can be ``listed'' as $\R = \{x_1, x_2, \ldots\} = \{x_n\}_{n \in \N}$. We will define a sequence of non-empty closed intervals $\{I_k\}_{k \in \N}$ such that $I_{k+1} \subseteq I_k$ and $x_k \notin I_k$ for all $k \in \N$. Let $I_0 \coloneq [0,1]$. Divide $I_0$ into three equal closed intervals. Then, at least one of these three intervals does not contain $x_1$. Choose such an interval and call it $I_1$. Divide $I_1$ into there equal closed intervals. Then, at least one of these three intervals does not contain $x_2$. Choose such an interval and call it $I_2$. Given $I_k$ for some $k \in \N$, divide $I_k$ into three equal closed intervals, then choose the interval that does not contain $x_{k+1}$ and call it $I_{k+1}$. By induction, we have $\{I_k\}_{k=1}^\infty$ where each $I_k$ is a (nonempty) closed interval, and $I_{k+1} \subseteq I_k$ for each $k \in \N$. By the \nameref{thm:nested-interval-property}, $\bigcap_{k \in \N} I_k$ is not empty, so there exists $x \in \R$ such that $x \in \bigcap_{k \in \N} I_k$. Since $x \in \R$, we have $x = x_n$ for some $n \in \N$ (by our supposition that $\R$ is countable). However, we constructed $I_n$ such that $x_n \notin I_n$, so $x_n \notin \bigcap_{k \in \N} I_k$. This contradiction renders our initial supposition false. Therefore, $\R$ is uncountable.
    \end{proof}
\end{thmbox}

\begin{thmbox}{Irrational Numbers are Uncountable}{}
    \begin{proof}
        Suppose for contradiction $\R \setminus \Q$ is countable. Then $\R = (\R \setminus \Q) \cup \Q$, so $\R$ is a countable union of countable sets, making $\R$ countable. This contradicts the fact that $\R$ is uncountable, so $\R \setminus \Q$ is uncountable.
    \end{proof}
\end{thmbox}

\section{Additional Remarks}

\begin{dfnbox}{Algebraic Number, Transcandental Number}{}
    If $\alpha$ is a root of the polynomial $a_nx^n + a_{n-1} x^{n-1} + \cdots + a_1 x + a_0 = 0$ where each $a_i \in \Z$, then $\alpha$ is called an \dfntxt{algebraic number}. If $\alpha$ is not algebraic, then we call it a \dfntxt{transcendental number}.
\end{dfnbox}
 The set of all algebraic numbers is countable, so ``most'' real numbers are transcendental.

\begin{itemize}
    \item Even though $\Q$ is dense in $\R$, there are ``more'' irrational numbers than rational numbers
    \item The set $\{\sqrt[n]{m} : n,m \in \N\}$ is countable, so ``most'' real numbers are not radicals.
    \item The set of algebraic numbers is countable, so ``most'' real numbers are transcendental.
    \item $\abs{\mathcal{P}(\N)} = \abs{\R}$, but the set $\{ A \subseteq \N : A\ \text{is finite} \}$ is countable.
    \item If there exists an injection $f : A \to B$, then we say $\abs{A} \leq \abs{B}$. If there exists an injection from $A$ to $B$, but there does not exist an injection from $B$ to $A$, then we say $\abs{A} < \abs{B}$.
\end{itemize}

\begin{thmbox}{Schroeder-Bernstein}{}
    If there exists an injection $f : A \to B$ and an injection $g : B \to A$, then there exists a bijection $h : A \to B$.
\end{thmbox}

\begin{thmbox}{Continuum Hypothesis}{}
    There is no cardinality between $\abs{\N}$ and $\abs{\R}$.
\end{thmbox}

In Zermelo-Fraenkel with Choice (ZFC) set theory, the continuum hypothesis cannot be proven to be true nor false. In 1938, Godel proved that the continuum hypothesis is consistent with ZFC. In 1963, Colen proved that the negation of the continuum hypothesis is also consisten with ZFC.

\begin{notebox}
    Just because you can write a description of a set does not mean that the set exists nor makes sense.
    \begin{itemize}
        \item For example, let $A \coloneq \left\{ \bigcup \{ B : B\ \text{is a set} \} \right\}$. Then $\mathcal{P}(A) \subseteq A$, so $\abs{\mathcal{P}(A)} \leq A < \abs{\mathcal{P}(A)}$.
        \item Another example: let $B \coloneq \{\text{all sets}\}$. Let $C \coloneq \{ A : A \notin A \}$. Is $C \in C$?
    \end{itemize}
\end{notebox}
